---
title: "Metaanalysis Assigment"
author: "Thuan Tran"
date: today
format:  
  html:
    theme: lumen
    toc: true
    toc_float: true
title-block-banner: true
execute:
  echo: fenced
---

```{r}
#| label: setup
#| include: false

library(meta)
library(metafor)
library(dplyr)
library(readxl)
```

## 1. Import the data

```{r}
#| echo: false

data <- read_excel("data/metaanalysis_data.xlsx")
head(data)
```
## 2. Experiment
```{r}
#| echo: false
data <- data |>
  mutate(
    Quality_CaseDef = ifelse(`Case definition adequate` == "*", 1, 0),
    Quality_Represent = ifelse(`Representativeness of cases` == "*", 1, 0),
    Quality_Controls = ifelse(`Selection of controls` == "*", 1, 0),
    Quality_ParentOpinion = ifelse(`Parental opinion` == "*", 1, 0),
    Quality_ComparableGroups = ifelse(grepl("\\*", `Comparability of both groups`), 1, 0),
    Quality_BehaviourDefined = ifelse(`Ascertainment of behaviour` == "*", 1, 0),
    Quality_SameMethod = rep(1, nrow(data)),  # Always fulfilled
    Quality_Nonresponse = ifelse(`Non‐response rate` == "*", 1, 0),
    Majority_Female = ifelse(`Female authors` > `Male authors`, 1, 0)
  )
```

### a. Combine the effects
```{r}
#| echo: false
m <- metacont(
  n.e = N_boys,
  mean.e = Mean_boys_play_male,
  sd.e = SD_boys_play_male,
  n.c = N_girls,
  mean.c = Mean_girls_play_male,
  sd.c = SD_girls_play_male,
  data = data,
  studlab = Study,
  sm = "SMD",
  method.tau = "REML",
  common = FALSE,
  random = TRUE
)
```

```{r}
#| echo: false
summary(m)
```
We can see that:

- 27 studies were included.

- the overall effect size (SMD) is 1.01 with a 95% confidence interval of [0.79; 1.23].

- This means boys played significantly more with male-typed toys than girls did.

- The p-value < 0.0001, showing a statistically significant difference.

- Heterogeneity is high:

   + I² = 71.1%, meaning much of the variation is due to differences between studies.

   + Tau² = 0.24, indicating moderate variability in effect sizes.


### b. Create a funnel plot
```{r}
#| echo: false
# Funnel plot to check publication bias
funnel(m, 
       contour = c(0.90, 0.95, 0.99),
       col.contour = c("darkblue", "blue", "lightblue"))
legend("topright", legend = c("p < 0.10", "p < 0.05", "p < 0.01"),
       fill = c("darkblue", "blue", "lightblue"))
```
We can observe that:

- the plot is slightly skewed, with more studies appearing on the right side. 

- This may suggest a possible publication bias: studies showing stronger effects (in favor of boys preferring male-typed toys) may be more likely to be published.

- However, the distribution is not extremely asymmetric, so the bias may be small.

### c. Check if methods / quality affect the results
```{r}
#| echo: false
# Meta-regression with method and quality variables
metareg(m, ~ `Neutral toys` + `Parent present` + Setting + Country +
           Quality_CaseDef + Quality_Represent + Quality_Controls +
           Quality_ParentOpinion + Quality_ComparableGroups +
           Quality_BehaviourDefined + Quality_SameMethod + Quality_Nonresponse)
```
- We ran a model to see if the way studies were done (methods and quality) affected the results.
- These things did make a difference:
  - Not using neutral toys → stronger results.
  - Study setting (home, lab, nursery) affected results.
  - Countries with more gender equality had smaller effects.
  - Studies that matched boys and girls well showed different results.
  - Asking parents about gender views made results stronger.
  - Reporting dropouts also affected results.
  - The model explained about 10% of the differences between studies.
  - There is still a lot of variation (I² = 72%) that is not explained.

*About the warnings:*

- One study was skipped because of missing data.

- Some variables were removed automatically because they didn’t help the model.

### d. Author Gender
```{r}
#| echo: false
# Create a new variable: more female than male authors
data <- data |>
  mutate(Majority_Female = ifelse(`Female authors` > `Male authors`, 1, 0))

# Meta-regression on author gender
metareg(m, ~ Majority_Female)
```
- We tested if the gender of the study authorsaffected the results.

- Specifically, we looked at whether studies with more female authors gave different results.

What we found:

- Studies with more female authors showed a smaller effect size.

- The difference was statistically significant (p = 0.011).

- This means author gender may influence the reported outcomes.

- The model explained about 22% of the variation between studies.
